%quant
\documentclass[11pt]{article} %Sets the default text size to 11pt and class to article.
\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}
\thispagestyle{empty}
%------------------------Dimensions--------------------------------------------
\topmargin=0.0in %length of margin at the top of the page (1 inch added by default)
\oddsidemargin=0.0in %length of margin on sides for odd pages
\evensidemargin=0in %length of margin on sides for even pages
\textwidth=6.5in %How wide you want your text to be
\marginparwidth=0.5in
\headheight=0pt %1in margins at top and bottom (1 inch is added to this value by default)
\headsep=0pt %Increase to increase white space in between headers and the top of the page
\textheight=9.0in %How tall the text body is allowed to be on each page
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\begin{document}
 \centerline{\Large \bf Shi Guo}
  \noindent \url{https://guoshi1984.github.io/} \hspace{80mm}  919-995-5612\\
    guoshi1984@hotmail.com \hspace{95mm}   Aliso Viejo, CA\\
\line(1,0){500}


\vspace{1mm}

\noindent{\large \bf{SUMMARY} }\\
\noindent 9 years of experience in designing, maintaining and testing large-scale C++/Java projects in both scientific research and industry.  \\
\noindent 5 years of experience of computational physics research focusing on quantum Monte Carlo simulation.\\
\noindent 5 years of experience in designing semiconductor testing program in c++/Java. Building and implementing test methodologies(including DC, digital, analog, radio frequency) for different kinds of semiconductor device \\
\noindent 4 years of experience of production data analysis using python with scipy, pandas library. 
Performing statistical tests including ANOVA, t test, F test. \\
\noindent [{\bf External Project}] Implemented and coded option pricing algorithms(analytical and Monte Carlo) based various different models(Black-Scholes, Heston, Stochastic jump diffusion). Paper and Code Link:  \url{https://guoshi1984.github.io/quant_lib} \\
\noindent Strong math background in linear algebra, numerical optimization, probability and measure theory, stochastic calculus.  \\
\noindent Strong physics background in quantum mechanics and solid state physics.

\vspace{2mm}

\noindent{\large \bf{SKILLS} }\\
\noindent 
Programming: C++, Java, Python scripting(Numpy), Matlab,  Linux(Bash Script, Vim) \\
Coursework and Certificate: Coursera Certificate in Machine Learning and Deep Learning, SAS(Regression and Modeling Certificate), Algorithms and Data Structure. %Subversion Control(SVN, GIT),  



\vspace{2mm}

\noindent{\large {\bf EXPERIENCE } }\\
\noindent{\bf Application Engineer}, Advantest America, Inc. July.2013 - Feb.2020\\
\noindent Developed and delivered system on chip(SoC) test programs for Advantest 93K automated tester platform.\\ 
\noindent{\bf Achievement Highlights} 
\begin{itemize}
  
    \item Developed test methods in C++/java for DC test, digital test, RF test.\\
 \vspace{-4mm}
    \item  Developed universal test method library for On-Die Parametric test on different TI DSP device, provide device characterization results to R\&D.\\ 
 \vspace{-4mm}
  \item Collaborated with customers to deliver test program for production.\\
 \vspace{-4mm}
  \item Provided on-site pre-sales support for different customers(TI, Broadcom) to trouble-shoot any tester hardware/software related problems.\\ 
  \vspace{-4mm}
 \item  Built data analysis tool using python(including scipy, pandas library) to do production data audit and statistical analysis(GR\&R).

  
 \end{itemize}

\noindent{\bf Research Assistant}, Department of Physics, North Carolina State University, Jun.2009-Jun.2013\\
\noindent Conducted computational physics research focusing on electronic structure using quantum Monte Carlo method, developed and maintained computational software code in C++ for 4 years.\\
\noindent {\bf Achievement Highlights}

\begin{itemize}
\item Conducted parallel-computing simulations on supercomputers using first-principles methods, including Hartree-Fock, Configuration Interaction, Density Functional Theory, Quantum Monte Carlo, etc.\\
\vspace{-4mm}
\item Conducted research using Quantum Monte Carlo method to evaluate different physical quantities in various kinds of system(atoms, molecules, solids).\\
\vspace{-4mm}
\item Maintained quantum Monte Carlo code under Linux-based cluster environment.\\
\vspace{-4mm}
\item Evaluated the dipole moment of weakly bonded ultracold molecules precisely using Quantum Monte Carlo method and provided the data with good quality to research funding agency.\\
\vspace{-4mm}
 \item Carried out numerical wavefunction optimizations(steep descent, quasi-Newton, etc.) to reduce the data fluctuation. \\
 \vspace{-4mm}
\item Introduced new mathematical models for electron spin and built into the existing algorithm to extend the features of the program to handle spin-dependent systems, designed every class and method independently using C++(10k lines). \\
 %\vspace{-4mm}
  %\item Fit numerical model to the calculation data using cross-validation fitting and present the result to research funding agency.
\vspace{-4mm}
  \item Performed the first successful calculation in literature on relativistic effects of two dimensional electron gas in semiconductor devices.\\
\end{itemize}



\vspace{2mm}

 
\noindent{\large \bf{EDUCATION} } \\
\vspace{-5mm}

\begin{itemize}
  \item {\bf Ph.D Physics} \hspace{100mm}   \\
    North Carolina State University, Raleigh, NC, USA, Dec.2013   \hspace*{10mm}GPA:3.63/4.0\\
\vspace{-4mm}
  \item {\bf B.S. Physics}, {\bf Minor in Economics}\\
    Shanghai Jiaotong University, Shanghai, China, Jun.2007\\
\vspace{-4mm}
\end{itemize}

\vspace{1mm}
\noindent{\large \bf{PUBLICATION}}\\
\vspace{-4mm}
\begin{itemize}
  \item {\bf Shi Guo}, Michal Bajdich, Lubos Mitas and Peter J. Reynolds;
Study of dipole moments of LiSr and KRb molecules by quantum Monte Carlo methods;
Molecular Physics Vol. 111, Iss. 12-13,2013
  \item Cody A. Melton, Minyi Zhu, {\bf Shi Guo}, Alberto Ambrosetti, Francesco Pederiva, and Lubos Mitas;
Spin-orbit interactions in electronic structure quantum Monte Carlo methods;
Phys. Rev. A 93, 042502
\end{itemize}
\vspace{-12mm}  

\end{document}
